---
title: "Fundamentals of Computing and Data Display"
subtitle: "Term paper template"
author: "Cameron Loats"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
    df_print: kable
references:
- id: Wickham2014
  title: Tidy Data
  author:
  - family: Wickham
    given: Hadley
  container-title: Journal of Statistical Software
  volume: 59
  issue: 10
  page: 1-23
  type: article-journal
  issued:
    year: 2014
- id: Baumer2017
  title: Modern Data Science with R
  author:
  - family: Baumer
    given: Benjamin S.
  - family: Kaplan
    given: Daniel T.
  - family: Horton
    given: Nicholas J.
  type: book
  publisher: Chapman \& Hall/CRC Press.
  issued:
    year: 2017
---

```{r, include = FALSE}
library(knitr)
library(tidyverse)
library(covidcast)
library(censusapi)
library(gridExtra)
```

## Introduction

This section outlines the research idea. We can also cite related work here [@Wickham2014; @Baumer2017].

Note that compiled term paper (the PDF) is supposed to be more text-centered than the RMarkdown documents we used in class, i.e. the text sections are more detailed and big or redundant code chunks can be hidden.

## Data

This section describes the data sources and the data gathering process.
The COVIDcast package from the Delphi group at Carnegie Mellon University contains data from 19 sources called signals within the package. Each of these signals represent different sources like the COVID-19 Trends and Impact Facebook survey, the CDC, and others. Within those sources there are many variables represented. I wanted to look at proportions of people with COVID-19, both from PCR test results and also estimates from the facebook survey. Beyond that I wanted to see the differences in the data for variables that showed potential exposure. For the first of those variables I wanted to look at estimated use of public transit. The thought here is that being in tight confined spaces, especially in the first year of the pandemic, would increase your risk of catching COVID-19.
The COVIDcast data was accessed using their API and the COVIDcast R package. With the package you select a data source, a signal within that data source, a time period, and a geographic location. For all of my variables I looked at both Prince George's and Montgomery counties. In addition I looked at the time period of October 1st to December 31st. I chose this time period because I wanted to see how the data compared in the period before vaccines were made available to the public. 
```{r}
cs_key <- "1448b92acdfc46410a020ee82b98af8fad7de580"
census_pg <- getCensus(name = "acs/acs5",
                    vintage = 2020, 
                    vars = c("NAME", "B01001_001E", "B06002_001E", "B19013_001E", "B19301_001E"), 
                    region = "county:033", 
                   regionin = "state:24",
                    key = cs_key)
census_pg <-
  census_pg %>%
  rename(pop = B01001_001E, age = B06002_001E, hh_income = B19013_001E, income = B19301_001E)
knitr::kable(head(census_pg), caption = "Census data for Prince George's County")
census_moco <- getCensus(name = "acs/acs5",
                    vintage = 2020, 
                    vars = c("NAME", "B01001_001E", "B06002_001E", "B19013_001E", "B19301_001E"), 
                    region = "county:031", 
                   regionin = "state:24",
                    key = cs_key)
census_moco <-
  census_moco %>%
  rename(pop = B01001_001E, age = B06002_001E, hh_income = B19013_001E, income = B19301_001E)
knitr::kable(head(census_moco), caption = "Census data for Montgomery County")
```

```{r}
# A code chunk that exemplifies the data gathering process
#code to find the FIPS codes for our counties of interest
pg_fips <- name_to_fips("Prince George's")
moco_fips <- name_to_fips("Montgomery", state = "MD")
#code to find the proportion of positive PCR tests for a county within a certain time period
pg_pcr_pos <- covidcast_signal(data_source = "covid-act-now", signal = "pcr_specimen_positivity_rate",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = pg_fips)
#knitr::kable(head(pg_pcr_pos),)
moco_pcr_pos <- covidcast_signal(data_source = "covid-act-now", signal = "pcr_specimen_positivity_rate",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = moco_fips)
#knitr::kable(head(moco_pcr_pos))

```

```{r, include = FALSE}
# Additional code chunks that repeat tasks or do basic things can be hidden
#code to find the estimated proportion of people with covid-like illnesses from the 
#COVID-19 Trends and Impacts survey "Estimated percentage of people with COVID-like illness"
pg_fb_pos <- covidcast_signal(data_source = "fb-survey", signal = "smoothed_wcli",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = pg_fips)
#knitr::kable(head(pg_fb_pos))
moco_fb_pos <- covidcast_signal(data_source = "fb-survey", signal = "smoothed_wcli",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = moco_fips)
#knitr::kable(head(moco_fb_pos))
#code looking at "Estimated percentage of respondents who “used public transit” in the past 24 hours"
pg_fb_bus <- covidcast_signal(data_source = "fb-survey", signal = "smoothed_wpublic_transit_1d",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = pg_fips)
#knitr::kable(head(pg_fb_bus))
moco_fb_bus <- covidcast_signal(data_source = "fb-survey", signal = "smoothed_wpublic_transit_1d",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = moco_fips)
#knitr::kable(head(moco_fb_bus))
#looking at "Estimated percentage of respondents who reported feeling “nervous, anxious, or on edge” for most or all of the past 5 days"
pg_fb_anxious <- covidcast_signal(data_source = "fb-survey", signal = "smoothed_wanxious_5d",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = pg_fips)
#knitr::kable(head(pg_fb_anxious))
moco_fb_anxious <- covidcast_signal(data_source = "fb-survey", signal = "smoothed_wanxious_5d",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = moco_fips)
#knitr::kable(head(moco_fb_anxious))
#looking at "Cumulative number of confirmed deaths due to COVID-19" from the JHU data
pg_jhu_deaths <- covidcast_signal(data_source = "jhu-csse", signal = "deaths_cumulative_num",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = pg_fips)
#knitr::kable(head(pg_jhu_deaths))
moco_jhu_deaths <- covidcast_signal(data_source = "jhu-csse", signal = "deaths_cumulative_num",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = moco_fips)
#knitr::kable(head(moco_jhu_deaths))
#Estimated percentage of people who wore a mask for most or all of the time while in public in the past 5 days; those not in public in the past 5 days are not counted.
pg_fb_mask <- covidcast_signal(data_source = "fb-survey", signal = "smoothed_wwearing_mask",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = pg_fips)
#knitr::kable(head(pg_fb_mask))
moco_fb_mask <- covidcast_signal(data_source = "fb-survey", signal = "smoothed_wwearing_mask",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = moco_fips)
#knitr::kable(head(moco_fb_mask))
#Estimated percentage of respondents who worked or went to school outside their home in the past 24 hours 
pg_fb_out <- covidcast_signal(data_source = "fb-survey", signal = "smoothed_wwork_outside_home_1d",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = pg_fips)
#knitr::kable(head(pg_fb_out))
moco_fb_out <- covidcast_signal(data_source = "fb-survey", signal = "smoothed_wwork_outside_home_1d",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = moco_fips)
#knitr::kable(head(moco_fb_out))
#Estimated percentage of respondents who report being very or somewhat worried about their “household’s finances for the next month” 
pg_fb_money <- covidcast_signal(data_source = "fb-survey", signal = "smoothed_wworried_finances",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = pg_fips)
#knitr::kable(head(pg_fb_money))
moco_fb_money <- covidcast_signal(data_source = "fb-survey", signal = "smoothed_wworried_finances",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = moco_fips)
#knitr::kable(head(moco_fb_money))
#Estimated percentage of new hospital admissions with COVID-associated diagnoses, based on claims data from health system partners, smoothed in time using a Gaussian linear smoother .  
pg_hhs_hospital <- covidcast_signal(data_source = "hospital-admissions", signal = "smoothed_covid19_from_claims",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = pg_fips)
#knitr::kable(head(pg_hhs_hospital))
moco_hhs_hospital <- covidcast_signal(data_source = "hospital-admissions", signal = "smoothed_covid19_from_claims",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = moco_fips)
#knitr::kable(head(moco_hhs_hospital))
```

## Results

This section presents the main results.

### Data exploration

The results section may have a data exploration part, but in general the structure here depends on the specific project.

```{r}
# What happens here depends on the specific project
#Data of PCR Positive Test Proportion
vis_pg_pcr_pos <- pg_pcr_pos %>% select(data_source ,signal, geo_value, time_value,value, sample_size)
knitr::kable(head(vis_pg_pcr_pos),caption = "Proportion of PCR Positive Tests for Prince George's County")
vis_moco_pcr_pos <- moco_pcr_pos %>% select(data_source ,signal, geo_value, time_value,value, sample_size)
knitr::kable(head(vis_moco_pcr_pos),caption = "Proportion of PCR Positive Tests for Montgomery County")
#Estimated proportion of individuals with COVID-19
vis_pg_fb_pos <- pg_fb_pos %>% select(data_source ,signal, geo_value, time_value,value, sample_size)
knitr::kable(head(vis_pg_fb_pos),caption = "Estimated proportion of individuals with COVID-19 for Prince George's County")
vis_moco_fb_pos <- moco_fb_pos %>% select(data_source ,signal, geo_value, time_value,value, sample_size)
knitr::kable(head(vis_moco_fb_pos),caption = "Estimated proportion of individuals with COVID-19 for Montgomery County")
#Estimated percentage of people who used public transit in the past day
vis_pg_fb_bus <- pg_fb_bus %>% select(data_source ,signal, geo_value, time_value,value, sample_size)
knitr::kable(head(vis_pg_fb_bus),caption = "Estimated percentage of people who used public transit in the past day for Prince George's County")
vis_moco_fb_bus <- moco_fb_bus %>% select(data_source ,signal, geo_value, time_value,value, sample_size)
knitr::kable(head(vis_moco_fb_bus),caption = "Estimated percentage of people who used public transit in the past day for Montgomery County")
#Estimated percentage of people who felt anxious within the past 5 days
vis_pg_fb_anxious <- pg_fb_anxious %>% select(data_source ,signal, geo_value, time_value,value, sample_size)
knitr::kable(head(vis_pg_fb_anxious),caption = "Estimated percentage of people who felt anxious within the past 5 days for Prince George's County")
vis_moco_fb_anxious <- moco_fb_anxious %>% select(data_source ,signal, geo_value, time_value,value, sample_size)
knitr::kable(head(vis_moco_fb_anxious),caption = "Estimated percentage of people who felt anxious within the past 5 days for Montgomery County")
#Estimated percentage of people who masked most of the time in the past 5 days
vis_pg_fb_mask <- pg_fb_mask %>% select(data_source ,signal, geo_value, time_value,value, sample_size)
knitr::kable(head(vis_pg_fb_mask),caption = "Estimated percentage of people who masked most of the time in the past 5 days for Prince George's County")
vis_moco_fb_mask <- moco_fb_mask %>% select(data_source ,signal, geo_value, time_value,value, sample_size)
knitr::kable(head(vis_moco_fb_mask),caption = "Estimated percentage of people who masked most of the time in the past 5 days for Montgomery County")
#Estimated percentage of respondents who worked or went to school outside their home in the past 24 hours 
vis_pg_fb_out <- pg_fb_out %>% select(data_source ,signal, geo_value, time_value,value, sample_size)
knitr::kable(head(vis_pg_fb_out),caption = "Estimated percentage of respondents who worked or went to school outside their home in the past 24 hours for Prince George's County")
vis_moco_fb_out <- moco_fb_out %>% select(data_source ,signal, geo_value, time_value,value, sample_size)
knitr::kable(head(vis_moco_fb_out),caption = "Estimated percentage of respondents who worked or went to school outside their home in the past 24 hours for Montgomery County")
#Estimated percentage of respondents who report being very or somewhat worried about their “household’s finances for the next month” 
vis_pg_fb_money <- pg_fb_money %>% select(data_source ,signal, geo_value, time_value,value, sample_size)
knitr::kable(head(vis_pg_fb_out),caption = "Estimated percentage of respondents who report being very or somewhat worried about their “household’s finances for the next month for Prince George's County")
vis_moco_fb_money <- moco_fb_money %>% select(data_source ,signal, geo_value, time_value,value, sample_size)
knitr::kable(head(vis_moco_fb_money),caption = "Estimated percentage of respondents who report being very or somewhat worried about their “household’s finances for the next month for Montgomery County")
#Estimated percentage of new hospital admissions with COVID-associated diagnoses, based on claims data from health system partners, smoothed in time using a Gaussian linear smoother .  
vis_pg_hhs_hospital <- pg_hhs_hospital %>% select(data_source ,signal, geo_value, time_value,value, sample_size)
knitr::kable(head(vis_pg_hhs_hospital),caption = "Estimated percentage of new hospital admissions with COVID-associated diagnoses for Prince George's County")
vis_moco_hhs_hospital <- moco_hhs_hospital %>% select(data_source ,signal, geo_value, time_value,value, sample_size)
knitr::kable(head(vis_moco_hhs_hospital),caption = "Estimated percentage of new hospital admissions with COVID-associated diagnoses for Montgomery County")
#looking at "Cumulative number of confirmed deaths due to COVID-19" from the JHU data
vis_pg_jhu_deaths <- pg_jhu_deaths %>% select(data_source ,signal, geo_value, time_value,value, sample_size)
knitr::kable(head(vis_pg_jhu_deaths),caption = "Cumulative number of confirmed deaths due to COVID-19 for Prince George's County")
vis_moco_jhu_deaths <- moco_jhu_deaths %>% select(data_source ,signal, geo_value, time_value,value, sample_size)
knitr::kable(head(vis_moco_jhu_deaths),caption = "Cumulative number of confirmed deaths due to COVID-19 for Montgomery County")
```

```{r}
# What happens here depends on the specific project
```

### Analysis

This section presents the main results, such as (for example) stats and graphs that show relationships, model results and/or clustering, PCA, etc.

```{r}
# What happens here depends on the specific project
#Graph of PCR positive proportion over time for Prince George's County vs Montgomery County
pos_pg <- pg_pcr_pos %>%
  group_by(time_value, value)
pos_moco <- moco_pcr_pos %>%
  group_by(time_value, value)
pos <- ggplot()+geom_line(data=pos_pg, aes(time_value,value), color="blue") + 
  geom_line(data=pos_moco, aes(time_value,value), color="black") + labs(title= "PCR Positive Test Proportion Over Time")
pos

```

```{r}
# What happens here depends on the specific project
fb_pg <- pg_fb_pos %>%
  group_by(time_value, value)
fb_moco <- moco_fb_pos %>%
  group_by(time_value, value)
pos_fb <- ggplot()+geom_line(data=fb_pg, aes(time_value,value), color="blue") + 
  geom_line(data=fb_moco, aes(time_value,value), color="black") + labs(title= "Estimated Proportion of People with COVID-like Illnesses")
pos_fb
```

```{r}
# What happens here depends on the specific project
bus_pg <- pg_fb_bus %>%
  group_by(time_value, value)
bus_moco <- moco_fb_bus %>%
  group_by(time_value, value)
bus <- ggplot()+geom_line(data=bus_pg, aes(time_value,value), color="blue") + 
  geom_line(data=bus_moco, aes(time_value,value), color="black") + labs(title= "Percentage of Respondents Who Used Public Transit in the Past 24 Hours")
bus
```
```{r}
anxious_pg <- pg_fb_anxious %>%
  group_by(time_value, value)
anxious_moco <- moco_fb_anxious %>%
  group_by(time_value, value)
anxious <- ggplot()+geom_line(data=anxious_pg, aes(time_value,value), color="blue") + 
  geom_line(data=anxious_moco, aes(time_value,value), color="black") + labs(title= "Estimated Percentage of Respondents Who Felt Anxious \nOver the Past 5 Days")
anxious
```
```{r}
mask_pg <- pg_fb_mask %>%
  group_by(time_value, value)
mask_moco <- moco_fb_mask %>%
  group_by(time_value, value)
mask <- ggplot()+geom_line(data=mask_pg, aes(time_value,value), color="blue") + 
  geom_line(data=mask_moco, aes(time_value,value), color="black") + labs(title= "Estimated Percentage of Respondents Who Masked for Most \nof the Past 5 Days")
mask
```
```{r}
out_pg <- pg_fb_out %>%
  group_by(time_value, value)
out_moco <- moco_fb_out %>%
  group_by(time_value, value)
out <- ggplot()+geom_line(data=out_pg, aes(time_value,value), color="blue") + 
  geom_line(data=out_moco, aes(time_value,value), color="black") + labs(title= "Est. Percentage of Respondents Who Worked or Went to School \nOutside their Home")
out
```
```{r}
money_pg <- pg_fb_money %>%
  group_by(time_value, value)
money_moco <- moco_fb_money %>%
  group_by(time_value, value)
money <- ggplot()+geom_line(data=money_pg, aes(time_value,value), color="blue") + 
  geom_line(data=money_moco, aes(time_value,value), color="black") + labs(title= "Est. Percentage of Respondents Who are Worried \nabout their Finaces Next Month")
money
```
```{r}
hospitals_pg <- pg_hhs_hospital %>%
  group_by(time_value, value)
hospitals_moco <- moco_hhs_hospital %>%
  group_by(time_value, value)
hospitals <- ggplot()+geom_line(data=hospitals_pg, aes(time_value,value), color="blue") + 
  geom_line(data=hospitals_moco, aes(time_value,value), color="black") + labs(title= "Est. Percentage of New Hospital Admissions with \nCOVID-Associated Diagnoses")
hospitals
```

```{r}
deaths_pg <- pg_jhu_deaths %>%
  group_by(time_value, value)
deaths_moco <- moco_jhu_deaths %>%
  group_by(time_value, value)
deaths <- ggplot()+geom_line(data=deaths_pg, aes(time_value,value), color="blue") + 
  geom_line(data=deaths_moco, aes(time_value,value), color="black") + labs(title= "Cumulative number of confirmed deaths due to COVID-19")
deaths
```

## Discussion

This section summarizes the results and may briefly outline advantages and limitations of the work presented.

## References
