---
title: "Fundamentals of Computing and Data Display"
subtitle: "Term paper template"
author: "Cameron Loats"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
    df_print: kable
references:
- id: Gravlee2020
  title: "Systemic racism, chronic health inequities, and COVID-19: A syndemic in the making?"
  author:
  - family: Gravlee
    given: Clarence C.
  container-title: "American journal of human biology: the official journal of the Human Biology Council"
  volume: 32
  issue: 5
  page: 
  type: article-journal
  issued:
    year: 2020
- id: Kim2020
  title: Social Vulnerability and Racial Inequality in COVID-19 Deaths in Chicago
  author:
  - family: Kim
    given: Sage J.
  - family: Bostwick
    given: Wendy
  container-title: Health Education & Behavior
  volume: 47
  issue: 4
  page: 
  type: article-journal
  issued:
    year: 2020
- id: Khazanchi2020
  title: Racism, Not Race, Drives Inequity Across the COVID-19 Continuum
  author:
  - family: Khazanchi
    given: Rohan
  - family: Evans
    given: Charlesnika T.
  - family: Marcelin
    given: Jasmine R.
  container-title: JAMA Netw Open
  volume: 3
  issue: 9
  page: 
  type: article-journal
  issued:
    year: 2020
---

```{r, include = FALSE}
library(knitr)
library(tidyverse)
library(covidcast)
library(censusapi)
library(gridExtra)
```

## Introduction

COVID-19 represents the first global pandemic of the 21st century. But despite being a global pandemic it did not affect every area equally. Not only were countries in the Global South hit very hard but certain areas of more developed countries also were hit harder than the rest of their country. Within the USA the areas hit hardest are those with higher minority populations and especially with higher Black populations [@Gravlee2020; @Kim2020; @Khazanchi2020]. This paper looks at the effect of COVID-19 on Prince George's County. Prince George's County was chosen for a few reasons. Firstly it is the most populated county in the country with a majority Black population. Secondly while not having the highest median household income of counties it does rank in the top 100 median household incomes. This should help control for the effect of income and healthcare. For a control county, Montgomery County was chosen. Montgomery County not only has a much smaller proportion of Black people but it is also geographically close to Prince George's County, with the two sharing a border. Having a county in the same state should help control for the effects of State-level governments on COVID-19.

The time period looked at for this paper is October 1st, 2020 through December 31st, 2021. This period was chosen as it represents the pandemic having existed for some months but before vaccinations were being administered. Not having to control for vaccine rollouts allows for simpler comparisons cross-county. In addition it was around this time period that COVID-19 was seen to be affecting areas with high minority populations much more than lower minority areas.

## Data 

The COVIDcast package from the Delphi group at Carnegie Mellon University contains data from 19 sources referred to as signals within the package. Each of these signals represent different sources like the COVID-19 Trends and Impact Facebook survey, the CDC, and others. Within those signals there are many variables represented. For data from the Facebook survey, the smoothed weighted version of a variable was always chosen so as to allow for cross-county comparisons more easily.

I wanted to look at proportions of people with COVID-19, both from PCR test results and also estimates from the Facebook survey. Beyond that I wanted to see the differences in the data for variables that showed potential exposure and also feelings during the pandemic. For the first of those variables I wanted to look at estimated use of public transit. The thought here is that being in tight confined spaces, especially in the first year of the pandemic, would increase your risk of catching COVID-19. How often people felt anxious was looked at from the Facebook survey. This measure was chosen to see if there's a difference in feelings of anxiety between the two counties. Use of masks was looked at as another variable to see potential exposure. Alongside that we also looked at percentage of individuals that had to either work or go to school outside their home. Lastly from the Facebook survey we looked at how many people were worried about their finances. The thought behind all of these is that Prince George's County should, as the county with the higher proportion of Black people and the lower median income, be the county where more people have to go out to work instead of working from home. With that they should use public transit more, be more anxious, and potentially more worried about finances. Masking should be similar between the two counties pre-vaccines but there is a possibility it will be more prevalent in Prince George's County if that county feels like it's being hit harder by COVID-19.

For more quantitative data I also looked at hospitalization rates and death rates. Hospitalization rates used data from electronic medical records and claims data about new admissions to the hospital for individuals with COVID-19. Death rates were taken from the Johns Hopkins University COVID-19 data, both cumulative and adjusted rates were gathered.

The COVIDcast data was accessed using their API and the COVIDcast R package. With the package you select a data source, a signal within that data source, a time period, and a geographic location. For all of my variables I looked at both Prince George's and Montgomery counties. In addition I looked at the time period of October 1st to December 31st.

The CensusAPI package provides a way to access the Census's API within R. For the purposes of this analysis Census data is used in a supplementary fashion. Population, age and household income were gathered for each county to see how different they were. Beyond that the White, Black and Latino population numbers were gathered. All of these values were put into an R object for statistical manipulation.

```{r}
# A code chunk that exemplifies the data gathering process
#code to find the FIPS codes for our counties of interest
pg_fips <- name_to_fips("Prince George's")
moco_fips <- name_to_fips("Montgomery", state = "MD")
#code to find the proportion of positive PCR tests for a county within a certain time period
pg_pcr_pos <- covidcast_signal(data_source = "covid-act-now", signal = "pcr_specimen_positivity_rate",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = pg_fips)
moco_pcr_pos <- covidcast_signal(data_source = "covid-act-now", signal = "pcr_specimen_positivity_rate",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = moco_fips)

```

```{r, include = FALSE}
# Additional code chunks that repeat tasks or do basic things can be hidden
#code to find the estimated proportion of people with covid-like illnesses from the 
#COVID-19 Trends and Impacts survey "Estimated percentage of people with COVID-like illness"
pg_fb_pos <- covidcast_signal(data_source = "fb-survey", signal = "smoothed_wcli",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = pg_fips)
moco_fb_pos <- covidcast_signal(data_source = "fb-survey", signal = "smoothed_wcli",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = moco_fips)
#code looking at "Estimated percentage of respondents who “used public transit” in the past 24 hours"
pg_fb_bus <- covidcast_signal(data_source = "fb-survey", signal = "smoothed_wpublic_transit_1d",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = pg_fips)
moco_fb_bus <- covidcast_signal(data_source = "fb-survey", signal = "smoothed_wpublic_transit_1d",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = moco_fips)
#looking at "Estimated percentage of respondents who reported feeling “nervous, anxious, or on edge” for most or all of the past 5 days"
pg_fb_anxious <- covidcast_signal(data_source = "fb-survey", signal = "smoothed_wanxious_5d",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = pg_fips)
moco_fb_anxious <- covidcast_signal(data_source = "fb-survey", signal = "smoothed_wanxious_5d",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = moco_fips)
#looking at "Cumulative number of confirmed deaths due to COVID-19" from the JHU data
pg_jhu_deaths <- covidcast_signal(data_source = "jhu-csse", signal = "deaths_cumulative_num",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = pg_fips)
moco_jhu_deaths <- covidcast_signal(data_source = "jhu-csse", signal = "deaths_cumulative_num",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = moco_fips)
#Estimated percentage of people who wore a mask for most or all of the time while in public in the past 5 days; those not in public in the past 5 days are not counted.
pg_fb_mask <- covidcast_signal(data_source = "fb-survey", signal = "smoothed_wwearing_mask",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = pg_fips)
moco_fb_mask <- covidcast_signal(data_source = "fb-survey", signal = "smoothed_wwearing_mask",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = moco_fips)
#Estimated percentage of respondents who worked or went to school outside their home in the past 24 hours 
pg_fb_out <- covidcast_signal(data_source = "fb-survey", signal = "smoothed_wwork_outside_home_1d",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = pg_fips)
moco_fb_out <- covidcast_signal(data_source = "fb-survey", signal = "smoothed_wwork_outside_home_1d",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = moco_fips)
#Estimated percentage of respondents who report being very or somewhat worried about their “household’s finances for the next month” 
pg_fb_money <- covidcast_signal(data_source = "fb-survey", signal = "smoothed_wworried_finances",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = pg_fips)
moco_fb_money <- covidcast_signal(data_source = "fb-survey", signal = "smoothed_wworried_finances",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = moco_fips)
#Estimated percentage of new hospital admissions with COVID-associated diagnoses, based on claims data from health system partners, smoothed in time using a Gaussian linear smoother .  
pg_hhs_hospital <- covidcast_signal(data_source = "hospital-admissions", signal = "smoothed_covid19_from_claims",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = pg_fips)
moco_hhs_hospital <- covidcast_signal(data_source = "hospital-admissions", signal = "smoothed_covid19_from_claims",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = moco_fips)
#code for cumulative deaths per 100,000
pg_jhu_prop <- covidcast_signal(data_source = "jhu-csse", signal = "deaths_cumulative_prop",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = pg_fips)
moco_jhu_prop <- covidcast_signal(data_source = "jhu-csse", signal = "deaths_cumulative_prop",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = moco_fips)
```

```{r}
cs_key <- "1448b92acdfc46410a020ee82b98af8fad7de580"
census_pg <- getCensus(name = "acs/acs5",
                    vintage = 2020, 
                    vars = c("NAME", "B01001_001E", "B06002_001E", "B19013_001E", "B19301_001E","B02001_002E", "B02001_003E", "B03001_003E"), 
                    region = "county:033", 
                   regionin = "state:24",
                    key = cs_key)
census_pg <-
  census_pg %>%
  rename(pop = B01001_001E, age = B06002_001E, hh_income = B19013_001E, income = B19301_001E, White = B02001_002E, Black = B02001_003E, Latino = B03001_003E)
knitr::kable(head(census_pg), caption = "Census data for Prince George's County")
census_moco <- getCensus(name = "acs/acs5",
                    vintage = 2020, 
                    vars = c("NAME", "B01001_001E", "B06002_001E", "B19013_001E", "B19301_001E","B02001_002E", "B02001_003E", "B03001_003E"), 
                    region = "county:031", 
                   regionin = "state:24",
                    key = cs_key)
census_moco <-
  census_moco %>%
  rename(pop = B01001_001E, age = B06002_001E, hh_income = B19013_001E, income = B19301_001E, White = B02001_002E, Black = B02001_003E, Latino = B03001_003E)
knitr::kable(head(census_moco), caption = "Census data for Montgomery County")
```

## Results

### Data exploration

Shown below is a table of the first six results from one of each of the four different data sources used from the COVIDcast package. The tables have been trimmed for ease of reading to include only the data source, the signal from within the source, the county FIPS code (geo_value), date (time_value), value being measured, and sample size. The values from covid-act-now are represented as proportions. Data from fb-survey and hospital-admissions is shown as a percentage. Lastly the data from jhu-csse is represented as individuals.

```{r}
# What happens here depends on the specific project
#Data of PCR Positive Test Proportion
vis_pg_pcr_pos <- pg_pcr_pos %>% select(data_source ,signal, geo_value, time_value,value, sample_size)
knitr::kable(head(vis_pg_pcr_pos),caption = "Proportion of PCR Positive Tests for Prince George's County")
vis_moco_pcr_pos <- moco_pcr_pos %>% select(data_source ,signal, geo_value, time_value,value, sample_size)
knitr::kable(head(vis_moco_pcr_pos),caption = "Proportion of PCR Positive Tests for Montgomery County")
#Estimated percentage of people who masked most of the time in the past 5 days
vis_pg_fb_mask <- pg_fb_mask %>% select(data_source ,signal, geo_value, time_value,value, sample_size)
knitr::kable(head(vis_pg_fb_mask),caption = "Estimated percentage of people who masked most of the time in the past 5 days for Prince George's County")
vis_moco_fb_mask <- moco_fb_mask %>% select(data_source ,signal, geo_value, time_value,value, sample_size)
knitr::kable(head(vis_moco_fb_mask),caption = "Estimated percentage of people who masked most of the time in the past 5 days for Montgomery County")
#Estimated percentage of new hospital admissions with COVID-associated diagnoses, based on claims data from health system partners, smoothed in time using a Gaussian linear smoother .  
vis_pg_hhs_hospital <- pg_hhs_hospital %>% select(data_source ,signal, geo_value, time_value,value, sample_size)
knitr::kable(head(vis_pg_hhs_hospital),caption = "Estimated percentage of new hospital admissions with COVID-associated diagnoses for Prince George's County")
vis_moco_hhs_hospital <- moco_hhs_hospital %>% select(data_source ,signal, geo_value, time_value,value, sample_size)
knitr::kable(head(vis_moco_hhs_hospital),caption = "Estimated percentage of new hospital admissions with COVID-associated diagnoses for Montgomery County")
#looking at "Cumulative number of confirmed deaths due to COVID-19" from the JHU data
vis_pg_jhu_deaths <- pg_jhu_deaths %>% select(data_source ,signal, geo_value, time_value,value, sample_size)
knitr::kable(head(vis_pg_jhu_deaths),caption = "Cumulative number of confirmed deaths due to COVID-19 for Prince George's County")
vis_moco_jhu_deaths <- moco_jhu_deaths %>% select(data_source ,signal, geo_value, time_value,value, sample_size)
knitr::kable(head(vis_moco_jhu_deaths),caption = "Cumulative number of confirmed deaths due to COVID-19 for Montgomery County")
```
```{r, include = FALSE}
#Estimated percentage of people who used public transit in the past day
vis_pg_fb_bus <- pg_fb_bus %>% select(data_source ,signal, geo_value, time_value,value, sample_size)
knitr::kable(head(vis_pg_fb_bus),caption = "Estimated percentage of people who used public transit in the past day for Prince George's County")
vis_moco_fb_bus <- moco_fb_bus %>% select(data_source ,signal, geo_value, time_value,value, sample_size)
knitr::kable(head(vis_moco_fb_bus),caption = "Estimated percentage of people who used public transit in the past day for Montgomery County")
#Estimated percentage of people who felt anxious within the past 5 days
vis_pg_fb_anxious <- pg_fb_anxious %>% select(data_source ,signal, geo_value, time_value,value, sample_size)
knitr::kable(head(vis_pg_fb_anxious),caption = "Estimated percentage of people who felt anxious within the past 5 days for Prince George's County")
vis_moco_fb_anxious <- moco_fb_anxious %>% select(data_source ,signal, geo_value, time_value,value, sample_size)
knitr::kable(head(vis_moco_fb_anxious),caption = "Estimated percentage of people who felt anxious within the past 5 days for Montgomery County")
#Estimated percentage of respondents who worked or went to school outside their home in the past 24 hours 
vis_pg_fb_out <- pg_fb_out %>% select(data_source ,signal, geo_value, time_value,value, sample_size)
knitr::kable(head(vis_pg_fb_out),caption = "Estimated percentage of respondents who worked or went to school outside their home in the past 24 hours for Prince George's County")
vis_moco_fb_out <- moco_fb_out %>% select(data_source ,signal, geo_value, time_value,value, sample_size)
knitr::kable(head(vis_moco_fb_out),caption = "Estimated percentage of respondents who worked or went to school outside their home in the past 24 hours for Montgomery County")
#Estimated percentage of respondents who report being very or somewhat worried about their “household’s finances for the next month” 
vis_pg_fb_money <- pg_fb_money %>% select(data_source ,signal, geo_value, time_value,value, sample_size)
knitr::kable(head(vis_pg_fb_out),caption = "Estimated percentage of respondents who report being very or somewhat worried about their “household’s finances for the next month for Prince George's County")
vis_moco_fb_money <- moco_fb_money %>% select(data_source ,signal, geo_value, time_value,value, sample_size)
knitr::kable(head(vis_moco_fb_money),caption = "Estimated percentage of respondents who report being very or somewhat worried about their “household’s finances for the next month for Montgomery County")
#Estimated proportion of individuals with COVID-19
vis_pg_fb_pos <- pg_fb_pos %>% select(data_source ,signal, geo_value, time_value,value, sample_size)
knitr::kable(head(vis_pg_fb_pos),caption = "Estimated proportion of individuals with COVID-19 for Prince George's County")
vis_moco_fb_pos <- moco_fb_pos %>% select(data_source ,signal, geo_value, time_value,value, sample_size)
knitr::kable(head(vis_moco_fb_pos),caption = "Estimated proportion of individuals with COVID-19 for Montgomery County")

vis_pg_jhu_prop <- pg_jhu_prop %>% select(data_source ,signal, geo_value, time_value,value, sample_size)
vis_moco_jhu_prop <- moco_jhu_prop %>% select(data_source ,signal, geo_value, time_value,value, sample_size)
```

```{r, include = FALSE}
combo_pcr_pos <- full_join(vis_pg_pcr_pos, vis_moco_pcr_pos, by= c("data_source","signal", "geo_value", "time_value", "value", "sample_size"))
combo_pcr_pos[combo_pcr_pos==24031] <- "Montgomery County"
combo_pcr_pos[combo_pcr_pos==24033] <- "Prince George's County"
combo_pcr_pos <- combo_pcr_pos %>%
  rename(County=geo_value)

combo_fb_pos <- full_join(vis_pg_fb_pos, vis_moco_fb_pos, by= c("data_source","signal", "geo_value", "time_value", "value", "sample_size"))
combo_fb_pos[combo_fb_pos==24031] <- "Montgomery County"
combo_fb_pos[combo_fb_pos==24033] <- "Prince George's County"
combo_fb_pos <- combo_fb_pos %>%
  rename(County=geo_value)

combo_fb_bus <- full_join(vis_pg_fb_bus, vis_moco_fb_bus, by= c("data_source","signal", "geo_value", "time_value", "value", "sample_size"))
combo_fb_bus[combo_fb_bus==24031] <- "Montgomery County"
combo_fb_bus[combo_fb_bus==24033] <- "Prince George's County"
combo_fb_bus <- combo_fb_bus %>%
  rename(County=geo_value)

combo_fb_anxious <- full_join(vis_pg_fb_anxious, vis_moco_fb_anxious, by= c("data_source","signal", "geo_value", "time_value", "value", "sample_size"))
combo_fb_anxious[combo_fb_anxious==24031] <- "Montgomery County"
combo_fb_anxious[combo_fb_anxious==24033] <- "Prince George's County"
combo_fb_anxious <- combo_fb_anxious %>%
  rename(County=geo_value)

combo_fb_mask <- full_join(vis_pg_fb_mask, vis_moco_fb_mask, by= c("data_source","signal", "geo_value", "time_value", "value", "sample_size"))
combo_fb_mask[combo_fb_mask==24031] <- "Montgomery County"
combo_fb_mask[combo_fb_mask==24033] <- "Prince George's County"
combo_fb_mask <- combo_fb_mask %>%
  rename(County=geo_value)

combo_fb_money <- full_join(vis_pg_fb_money, vis_moco_fb_money, by= c("data_source","signal", "geo_value", "time_value", "value", "sample_size"))
combo_fb_money[combo_fb_money==24031] <- "Montgomery County"
combo_fb_money[combo_fb_money==24033] <- "Prince George's County"
combo_fb_money <- combo_fb_money %>%
  rename(County=geo_value)

combo_fb_out <- full_join(vis_pg_fb_out, vis_moco_fb_out, by= c("data_source","signal", "geo_value", "time_value", "value", "sample_size"))
combo_fb_out[combo_fb_out==24031] <- "Montgomery County"
combo_fb_out[combo_fb_out==24033] <- "Prince George's County"
combo_fb_out <- combo_fb_out %>%
  rename(County=geo_value)

combo_jhu_deaths <- full_join(vis_pg_jhu_deaths, vis_moco_jhu_deaths, by= c("data_source","signal", "geo_value", "time_value", "value", "sample_size"))
combo_jhu_deaths[combo_jhu_deaths==24031] <- "Montgomery County"
combo_jhu_deaths[combo_jhu_deaths==24033] <- "Prince George's County"
combo_jhu_deaths <- combo_jhu_deaths %>%
  rename(County=geo_value)

combo_jhu_prop <- full_join(vis_pg_jhu_prop, vis_moco_jhu_prop, by= c("data_source","signal", "geo_value", "time_value", "value", "sample_size"))
combo_jhu_prop[combo_jhu_prop==24031] <- "Montgomery County"
combo_jhu_prop[combo_jhu_prop==24033] <- "Prince George's County"
combo_jhu_prop <- combo_jhu_prop %>%
  rename(County=geo_value)

combo_hhs_hospital <- full_join(vis_pg_hhs_hospital, vis_moco_hhs_hospital, by= c("data_source","signal", "geo_value", "time_value", "value", "sample_size"))
combo_hhs_hospital[combo_hhs_hospital==24031] <- "Montgomery County"
combo_hhs_hospital[combo_hhs_hospital==24033] <- "Prince George's County"
combo_hhs_hospital <- combo_hhs_hospital %>%
  rename(County=geo_value)
```

The Census ACS 2020 data provides data on racial makeup of each of our counties. For this study we looked at White, Black and Latino. In order to see the difference in racial proportions across counties the data needed to be combined and reshaped. After this reshaping we can see in the bar graph below that the proportion of Latino population is similar between the counties. But the relative proportions of Black and White populations differs significantly.

```{r}
combo_pg_census <- census_pg %>% 
  select(NAME,pop,White,Black,Latino) %>%
  mutate(prop_white=White/pop, 
       prop_black=Black/pop,
       prop_latino=Latino/pop)
combo_moco_census <-census_moco %>%
  select(NAME,pop,White,Black,Latino) %>%
    mutate(prop_white=White/pop, 
       prop_black=Black/pop,
       prop_latino=Latino/pop) 
combo_census <- full_join(combo_pg_census, combo_moco_census, by = c("NAME","pop","White","Black", "Latino","prop_white", "prop_black", "prop_latino") )
combo_census_race <- combo_census %>%
  select(NAME, prop_white, prop_black, prop_latino) %>%
  gather(key="race", value="proportion", 2:4) %>%
  rename(County=NAME)
combo_census_race
census_race <- ggplot() + geom_bar(data=combo_census_race, aes(race, proportion, fill=County), stat="identity", position="dodge") 
census_race
```


A chi-squared test was performed for each pair of proportions to test equality. All three were statistically significantly not equal but the Latino difference was much less extreme. The proportions there were approximately 18.8% for Prince George's County and 19.5% for Montgomery County. On the other hand the proportion of Prince George's County that is Black is approximately 62.2% and for Montgomery County it is only 18.5%. Such a difference could very well be an explanation for why COVID-19 impacts Prince George's County differently than Montgomery County.

```{r}
# What happens here depends on the specific project
#Chi-squared test to determine if the difference between the proportions of the Black population in Prince George's and Montgomery County is statistically significant
sig <- prop.test(x = c(census_pg$Black, census_moco$Black), n = c(census_pg$pop, census_moco$pop))
sig
sig2 <- prop.test(x = c(census_pg$Latino, census_moco$Latino), n = c(census_pg$pop, census_moco$pop))
sig2
sig3 <- prop.test(x = c(census_pg$White, census_moco$White), n = c(census_pg$pop, census_moco$pop))
sig3
```
The data from the Census needs to be adjusted to proportions in order to properly compare it across counties and the following code does that.
```{r}
census_pg_prop <- census_pg %>%
  select(pop,White,Black,Latino) %>%
mutate(prop_white=White/pop, 
       prop_black=Black/pop,
       prop_latino=Latino/pop)
census_moco_prop <- census_moco %>%
  select(pop,White,Black,Latino) %>%
mutate(prop_white=White/pop, 
       prop_black=Black/pop,
       prop_latino=Latino/pop)
```


### Analysis

For all of the gathered variables I looked at the rate of them over time compared between the two counties. The Facebook survey data as a whole tends to look more erratic than the other data sources. This is most likely a result of how the survey was conducted and the estimation process used for the data.

First we have the positive proportion of PCR tests for both Prince George's County and Montgomery County. At all times the Prince George's line is higher than the Montgomery line. In addition, two of the times we see a spike in the Montgomery line, mid-October and mid-November, we see a correlating larger spike in the Prince George's line. There is a spike in early December for Montgomery that is not reciprocated by Prince George's but this could be a result of the very large spike for Prince George's in late November leading to under-testing in early to mid December. Throughout the time period we see Prince George's having a higher proporiton of positive PCR tests, which should correlate with a higher rate of COVID-19.

```{r}
#Graph of PCR positive proportion over time for Prince George's County vs Montgomery County
pos <- ggplot(data=combo_pcr_pos, aes(time_value,value))+geom_line(data=combo_pcr_pos, aes(color=County)) +  labs(title= "PCR Positive Test Proportion Over Time")
pos

```
In comparison with the PCR proportion, the estimated proportion of people with COVID-like illnesses from the Facebook survey shows the two counties closer together for large section of the three month period. We do still see areas where Prince George's county is higher, early to mid October and both early and mid December, but outside of those the lines are much closer to each other.

```{r}
pos_fb <- ggplot(data=combo_fb_pos, aes(time_value,value))+geom_line(data=combo_fb_pos, aes(color=County)) + labs(title= "Estimated Proportion of People with COVID-like Illnesses")
pos_fb
```
With percentage of people who used public transit in the past day we see a graph where neither county is consistently on top. At times residents in Prince George's county look to ride public transit more and at other times residents in Montgomery county look to ride it more. As such this does not appear to be a good metric to judge the effects of COVID-19 by.

```{r}
bus <- ggplot(data=combo_fb_bus, aes(time_value,value))+geom_line(data=combo_fb_bus, aes(color=County)) + labs(title= "Percentage of Respondents Who Used Public Transit in the Past 24 Hours")
bus
```
The variable looking at the percentage of respondents who felt anxious over the past 5 days differs from my hypothesis. I had assumed that we would see higher feelings of anxiety for those in Prince George's county due to lower median household income and the effects of systemic racism [@Gravlee2020]. Instead we see residents of Montgomery county showcasing higher feelings of anxiety relatively consistently. 

```{r}
anxious <- ggplot(data=combo_fb_anxious, aes(time_value,value))+geom_line(data=combo_fb_anxious, aes(color=County)) + labs(title= "Estimated Percentage of Respondents Who Felt Anxious \nOver the Past 5 Days")
anxious
```
Across both counties we see high masking percentages. At various points both counties are masking more than the other but throughout the time period both are masking at an estimated 91% or better. 

```{r}
mask <- ggplot(data=combo_fb_mask, aes(time_value,value))+geom_line(data=combo_fb_mask, aes(color=County)) + labs(title= "Estimated Percentage of Respondents Who Masked for Most \nof the Past 5 Days")
mask
```
Over time both counties have a smaller percentage of residents working or going to school outside their home. But within that trend Prince George's county residents were outside their home for work or school more than Montgomery especially within the period of mid-November to mid-December.

```{r}
out <- ggplot(data=combo_fb_out, aes(time_value,value))+geom_line(data=combo_fb_out, aes(color=County)) + labs(title= "Est. Percentage of Respondents Who Worked or Went to School \nOutside their Home")
out
```
Residents of Prince George's county were more worried about their finances than those in Montgomery county. The spikes seem to correlate to the spikes shown in PCR positive tests as well for Prince George's county.

```{r}
money <- ggplot(data=combo_fb_money, aes(time_value,value))+geom_line(data=combo_fb_money, aes(color=County)) + labs(title= "Est. Percentage of Respondents Who are Worried \nabout their Finaces Next Month")
money
```
Both counties experience an increase in new hospital admissions over the time period, starting at approximately 5% at the start and between 11% to 16% towards the end. Within the time period we can see two major spikes for Prince George's county, in late-October and mid-December, with only the later spike being represented in Montgomery county as well.

```{r}
hospitals <- ggplot(data=combo_hhs_hospital, aes(time_value,value))+geom_line(data=combo_hhs_hospital, aes(color=County)) + labs(title= "Est. Percentage of New Hospital Admissions with \nCOVID-Associated Diagnoses")
hospitals
```
For cumulative death rates, both counties follow roughly the same slope trajectory with Montgomery county having a consistently higher cumulative death toll throughout the time period. 
```{r}
deaths <- ggplot(data=combo_jhu_deaths, aes(time_value,value))+geom_line(data=combo_jhu_deaths, aes(color=County)) + labs(title= "Cumulative number of confirmed deaths due to COVID-19")
deaths
```
When adjusted to per 100,000 people we see that Prince George's county has the higher death rate throughout the time period.

```{r}
prop <- ggplot(data=combo_jhu_prop, aes(time_value,value))+geom_line(data=combo_jhu_prop, aes(color=County)) + labs(title= "Cumulative number of confirmed deaths due to COVID-19,\n per 100,000 population")
prop
```

## Discussion

For the time period of October 1st, 2020 through December 31st, 2020 we see the residents of Prince George's county being impacted more by COVID-19 then those in Montgomery county. Positive test rates, hospitalization and adjusted death rates were all higher for those in Prince George's county. One possible explanation is that Prince George's county has a higher proportion of Black residents. 

A limitation of this paper is that the data used from the Facebook survey does not provide exactly what is needed in order to make anything stronger than inferences. While the residents in Prince George's county were more worried about finances consistently than those in Montgomery, for every other metric looked at the differences were either much smaller or varied between which county had a higher percentage. 

Code is available in a git-hub repo at https://github.com/cloats1/727_Final_Project 

## References
