---
title: "Fundamentals of Computing and Data Display"
subtitle: "Term paper template"
author: "Cameron Loats"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
    df_print: kable
references:
- id: Wickham2014
  title: Tidy Data
  author:
  - family: Wickham
    given: Hadley
  container-title: Journal of Statistical Software
  volume: 59
  issue: 10
  page: 1-23
  type: article-journal
  issued:
    year: 2014
- id: Baumer2017
  title: Modern Data Science with R
  author:
  - family: Baumer
    given: Benjamin S.
  - family: Kaplan
    given: Daniel T.
  - family: Horton
    given: Nicholas J.
  type: book
  publisher: Chapman \& Hall/CRC Press.
  issued:
    year: 2017
---

```{r, include = FALSE}
library(knitr)
library(tidyverse)
library(covidcast)
library(censusapi)
library(gridExtra)
```

## Introduction

This section outlines the research idea. We can also cite related work here [@Wickham2014; @Baumer2017].

Note that compiled term paper (the PDF) is supposed to be more text-centered than the RMarkdown documents we used in class, i.e. the text sections are more detailed and big or redundant code chunks can be hidden.

## Data

This section describes the data sources and the data gathering process.
The COVIDcast package from the Delphi group at Carnegie Mellon University contains data from 19 sources called signals within the package. Each of these signals represent different sources like the COVID-19 Trends and Impact Facebook survey, the CDC, and others. Within those sources there are many variables represented. I wanted to look at proportions of people with COVID-19, both from PCR test results and also estimates from the facebook survey. Beyond that I wanted to see the differences in the data for variables that showed potential exposure. For the first of those variables I wanted to look at estimated use of public transit. The thought here is that being in tight confined spaces, especially in the first year of the pandemic, would increase your risk of catching COVID-19.
```{r}
#census code chunk that is failing
cs_key <- "1448b92acdfc46410a020ee82b98af8fad7de580"
census_pg <- getCensus(name = "acs/acs5",
                    vintage = 2020, 
                    vars = c("NAME", "B01001_001E", "B06002_001E", "B19013_001E", "B19301_001E"), 
                   # region = "county:*", 
                    region = "state:24+county:033",
                    key = cs_key)
head(census_pg)
```

```{r}
# A code chunk that exemplifies the data gathering process
#code to find the FIPS codes for our counties of interest
pg_fips <- name_to_fips("Prince George's")
moco_fips <- name_to_fips("Montgomery", state = "MD")
#code to find the proportion of positive PCR tests for a county within a certain time period
pg_pcr_pos <- covidcast_signal(data_source = "covid-act-now", signal = "pcr_specimen_positivity_rate",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = pg_fips)
knitr::kable(head(pg_pcr_pos))
moco_pcr_pos <- covidcast_signal(data_source = "covid-act-now", signal = "pcr_specimen_positivity_rate",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = moco_fips)
knitr::kable(head(moco_pcr_pos))
#code to find the estimated proportion of people with covid-like illnesses from the 
#COVID-19 Trends and Impacts survey "Estimated percentage of people with COVID-like illness"
pg_fb_pos <- covidcast_signal(data_source = "fb-survey", signal = "smoothed_wcli",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = pg_fips)
knitr::kable(head(pg_fb_pos))
moco_fb_pos <- covidcast_signal(data_source = "fb-survey", signal = "smoothed_wcli",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = moco_fips)
knitr::kable(head(moco_fb_pos))
#code looking at "Estimated percentage of respondents who “used public transit” in the past 24 hours"
pg_fb_bus <- covidcast_signal(data_source = "fb-survey", signal = "smoothed_wpublic_transit_1d",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = pg_fips)
knitr::kable(head(pg_fb_bus))
moco_fb_bus <- covidcast_signal(data_source = "fb-survey", signal = "smoothed_wpublic_transit_1d",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = moco_fips)
knitr::kable(head(moco_fb_bus))
#looking at "Estimated percentage of respondents who reported feeling “nervous, anxious, or on edge” for most or all of the past 5 days"
pg_fb_anxious <- covidcast_signal(data_source = "fb-survey", signal = "smoothed_wanxious_5d",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = pg_fips)
knitr::kable(head(pg_fb_anxious))
moco_fb_anxious <- covidcast_signal(data_source = "fb-survey", signal = "smoothed_wanxious_5d",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = moco_fips)
knitr::kable(head(moco_fb_anxious))
#looking at "Cumulative number of confirmed deaths due to COVID-19" from the JHU data
pg_jhu_deaths <- covidcast_signal(data_source = "jhu-csse", signal = "deaths_cumulative_num",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = pg_fips)
knitr::kable(head(pg_jhu_deaths))
moco_jhu_deaths <- covidcast_signal(data_source = "jhu-csse", signal = "deaths_cumulative_num",
                        start_day = "2020-10-01", end_day = "2020-12-31",
                        geo_type = "county", geo_value = moco_fips)
knitr::kable(head(moco_jhu_deaths))
```

```{r, include = FALSE}
# Additional code chunks that repeat tasks or do basic things can be hidden
```

## Results

This section presents the main results.

### Data exploration

The results section may have a data exploration part, but in general the structure here depends on the specific project.

```{r}
# What happens here depends on the specific project
```

```{r}
# What happens here depends on the specific project
```

### Analysis

This section presents the main results, such as (for example) stats and graphs that show relationships, model results and/or clustering, PCA, etc.

```{r}
# What happens here depends on the specific project
#Graph of PCR positive proportion over time for Prince George's County vs Montgomery County
pos_pg <- pg_pcr_pos %>%
  group_by(time_value, value)
pos_moco <- moco_pcr_pos %>%
  group_by(time_value, value)
pos <- ggplot()+geom_line(data=pos_pg, aes(time_value,value), color="blue") + 
  geom_line(data=pos_moco, aes(time_value,value), color="black")
pos

```

```{r}
# What happens here depends on the specific project
fb_pg <- pg_fb_pos %>%
  group_by(time_value, value)
fb_moco <- moco_fb_pos %>%
  group_by(time_value, value)
pos_fb <- ggplot()+geom_line(data=fb_pg, aes(time_value,value), color="blue") + 
  geom_line(data=fb_moco, aes(time_value,value), color="black")
pos_fb
```

```{r}
# What happens here depends on the specific project
```

## Discussion

This section summarizes the results and may briefly outline advantages and limitations of the work presented.

## References
